---
title: "Layers Blogpost Visualizations"
author: "Adelaide Robinson"
date: "2023-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in data

```{r}
library(googlesheets4)
library(tidyverse)
library(visNetwork)
library(htmlwidgets)

#read in the layers sheet (internal OHI spreadsheet used for keeping track of layers each year)
layers <- read_sheet("https://docs.google.com/spreadsheets/d/1hNzPZsTy_MpEcPg2aPJQfgCRO9njN3-qVNQ_4lzryng/edit?pli=1#gid=0") |> janitor::clean_names()

#read in the goal image icons from github
ao_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/AO.png?raw=true"
bd_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/BD.png?raw=true"
cp_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/CP.png?raw=true"
cs_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/CS.png?raw=true"
cw_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/CW.png?raw=true"
fp_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/FP.png?raw=true"
le_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/LE.png?raw=true"
np_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/NP.png?raw=true"
sp_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/SP.png?raw=true"
tr_image <- "https://github.com/OHI-Science/OHI-website/blob/main/content/images/goal_icons_old/TR.png?raw=true"
```

# General Data Wrangling

```{r}
# Define the words we want to remove
#Only considering if they affect a goal, not including dimensions
words_to_remove <- c("status", "trend", "resilience", "pressure")

# Use the paste function to create a regular expression pattern
pattern_to_remove <- paste(words_to_remove, collapse = "|")

# Apply str_remove with the pattern to the targets column
layers_intermediate <- layers %>%
  mutate(goals = str_remove_all(targets, pattern_to_remove)) |> 
  mutate(goals = str_replace_all(goals, "\\s+", " ")) |>  # Trim leading and trailing spaces
  separate_rows(goals, sep = " ") |> 
  filter(!is.na(goals)) |> 
  filter(!goals == "")

#clean up goals
layers_clean <- layers_intermediate |>
  filter(!goals %in% c("spatial", "weighting", "NA")) |> #remove NAs and static layers that don't impact any goals
  filter(!layer_name == "Food provision weights") |> 
  mutate(layer_id = as.numeric(factor(layer_name, levels = unique(layer_name)))-1) |> #visnetwork package requires ids to start at 0
  mutate(goal_id = as.numeric(factor(goals, levels = unique(goals))) + (length(unique(layer_id))-1)) #visnetwork package requires sequential ids that start at 0

#define unique layers
layer_ids <- layers_clean |>
  group_by(layer_id, layer_name) |> 
  summarize(n = n()) |> 
  select(label = layer_name, id = layer_id)

#define unique goals
goal_id <- layers_clean |> 
  group_by(goal_id, goals) |> 
  summarise(n = n()) |> 
  select(label = goals, id = goal_id)
```

# Card Visualization

Creating a simple visualization for the card on the website.

```{r}
#crate a data frame with all the elements we need
nodes <- data.frame(label = c("Raw Data", "Data Prep", "Data Layer", "Goal"),
                       id = c(0,1,2,3), 
                       group = c("Raw Data", "Data Prep", "Data Layer", "Goal"))
#define the connections
edges <- data.frame(from = c(0, 1,2),
                    to = c(1,2,3)) |> mutate(weight = 1)

visNetwork(nodes,edges) |>
  visHierarchicalLayout(
  direction = "LR", sortMethod = "directed"
) |> 
  visEdges(arrows = 'to') |> #add arrows
  visGroups(groupname = "Raw Data",
            color = "#fcbc6c",
            size =40,
            borders = 0) |> 
  visGroups(groupname = "Data Prep",
            color = "#f07446",
            size = 40) |> 
  visGroups(groupname = "Data Layer",
            color = "#377cbf",
            size = 40) |> 
  visGroups(groupname = "Goal",
            color = "#a7354e",
            size = 40)
```

# Mariculture Visualization

This code is used to make a visualization that shows the data preparation and incorporation process for the mariculture data prep.

```{r}
#filter to just the data layers that are created in the mariculture data prep
layers_mar <- layers_intermediate |>
  filter(layer_name %in% c("Genetic escapes", 	
"Mariculture sustainability score", 	
"Mariculture harvest")) |> 
  mutate(layer_id = as.numeric(factor(layer_name, levels = unique(layer_name)))-1) |> #create ids
  mutate(goal_id = as.numeric(factor(goals, levels = unique(goals))) + (length(unique(layer_id))-1)) #create ids

#create the edges data frame, used to create the lines between nodes
edges_mar <- layers_mar |>
  mutate(weight = 1) |> 
  select(from = layer_id, to = goal_id, weight) |> 
  mutate(from = as.numeric(from), to = as.numeric(to))

#create data frame with layer ids
layer_ids_mar <- layers_mar |> 
  group_by(layer_id, layer_name) |> 
  summarize(n = n()) |> 
  select(label = layer_name, id = layer_id)

#create data frame with ids for each goal
goal_id_mar <- layers_mar |> 
  group_by(goal_id, goals) |> 
  summarise(n = n()) |> 
  select(label = goals, id = goal_id)

#combine
nodes_mar <- rbind(layer_ids_mar, goal_id_mar) |> 
  mutate(id = as.numeric(id))

#create a dataframe of the raw data used and ids
data_raw <- data.frame(label= c("Monterey Bay Aquarium Seafood Watch Recommendations", "FAO Global Aquaculture Production", "Mar_dataprep.RMD"), id = c(8,9,10))

#combine the data frames
nodes_mar <- rbind(nodes_mar,data_raw) |> 
  mutate(group = case_when(
    label %in% c("MAR", "ECO", "FIS", "LIV", "SPP")~ "Goal",
    label %in% c("Mariculture harvest","Mariculture sustainability score",
                  "Genetic escapes") ~ "Data Layer",
    label %in% c("Monterey Bay Aquarium Seafood Watch Recommendations",
                "FAO Global Aquaculture Production") ~ "Raw Data",
    label == "Mar_dataprep.RMD" ~ "Data Prep")) |> #create groups 
  mutate(label = case_when(label =="MAR" ~"Mariculture",
                           label == "ECO"~ "Economies",
                           label == "FIS"~"Fisheries",
                           label =="LIV" ~"Livelihoods",
                          label == "SPP"~ "Species Condition",
                          TRUE ~label)) #add longer labels for subgoals

#order the dataframe so that the legend is in a logical order
nodes_mar <- nodes_mar[order(
  factor(nodes_mar$group, levels = c("Raw Data", "Data Prep", "Data Layer", "Goal"))
), ]

#create the data frame with connections
edges_new <- data.frame(from = c(8,9, 10, 10, 10), to = c(10,10, 0, 1, 2)) |> 
  mutate(weight = 1)

#combine and finish edges data frame
edges_final <- edges_new |> rbind(edges_mar) 

#create the visualization
visNetwork(nodes_mar, edges_final) |> 
   visHierarchicalLayout(
  direction = "LR", sortMethod = "directed"
) |> #make it a hierarchical layour
   visEdges(arrows = 'to') |> #add arrows
  visNodes(font = list(size = 13)) |> #update font size
  visGroups(
    groupname = "Raw Data",
    color = "#fcbc6c",
    size = 22,
    shape = "dot") |> #set size and color for raw data group
  visGroups(
    groupname = "Data Prep",
    color = "#f07446",
    size = 22,
    shape = "dot") |> #set size and color for data prep group
  visGroups(
    groupname = "Data Layer",
    color = "#377cbf",
    size = 22,
    shape = "dot") |> #set size and color for data layer group
   visGroups(
    groupname = "Goal",
    color = "#a7354e",
    size = 22,
    shape = "dot")  |>  #set size and color for goal layer group
  visLegend(useGroups = TRUE, 
            position = "right",
            width = .3,
            zoom = FALSE) |> #add a legend
  visOptions(width = "145%",
             height = "84%") |> #adjust height and width
    visPhysics(solver = "hierarchicalRepulsion", 
            hierarchicalRepulsion = list(nodeDistance = 150,
                                         springLength = 150)
    ) |>  visInteraction(zoomView = FALSE) #turn off zoom

#save to repository
saveWidget(mar_plot, "/Users/adelheid/Documents/Work/NCEAS/repositories/OHI-website/content/images/layers_blog/mar_connections.html")
```

# Layers Visualization

This code creates a visualization which shows the connections between all layers included in OHI and the goals which the impact.

```{r}
#make the nodes df with ids for all layers and goals
nodes <- rbind(layer_ids, goal_id) |> 
  mutate(id = as.numeric(id))|> 
  mutate(title = case_when(label == "AO" ~ "Artisinal Opportunity",
                           label == "CW" ~ "Clean Water",
                           label == "CP" ~ "Coastal Protection",
                           label == "CS" ~ "Carbon Storage",
                           label == "ECO" ~ "Economies",
                           label == "HAB" ~ "Habitat",
                           label == "ICO"~ "Iconic species",
                           label == "NP"~ "Natural Products",
                           label == "SPP" ~"Species Condition",
                           label == "LIV"~ "Livelihoods",
                           label == "LSP" ~ "Lasting Special Places",
                           label == "MAR" ~ "Mariculture",
                           label == "TR" ~ "Tourism and Recreation",
                           label =="FIS" ~ "Fisheries",
                           TRUE ~ label)) #use long names for labels

#create the edgeds df with the lines connecting layers and goals 
edges <- layers_clean |>
  mutate(weight = 1) |> 
  select(from = layer_id, to = goal_id, weight) |> 
  mutate(from = as.numeric(from), to = as.numeric(to)) |> 
  distinct()

#add additional features to the nodes data frame
nodes <- nodes %>%
  mutate(group = case_when(id > 96 ~ "goals",
                           TRUE ~ "layers")) |> #separate into goals and layers for grouping
  mutate(image = case_when(label == "AO" ~ ao_image, 
                           label == "CW" ~ cw_image,
                           label == "CP" ~ cp_image,
                           label == "CS" ~ cs_image,
                           label == "ECO" ~ le_image,
                           label == "HAB" ~ bd_image,
                           label == "ICO"~ sp_image,
                           label == "NP"~ np_image,
                           label == "SPP" ~bd_image,
                           label == "LIV"~ le_image,
                           label == "LSP" ~ sp_image,
                           label == "MAR" ~fp_image,
                           label == "TR" ~ tr_image,
                           label =="FIS" ~ fp_image,
                           TRUE ~ NA)) |> #add goal icon images
  mutate(goal = case_when( 
                           label == "ECO" ~ "Coastal Livelihoods and Economies",
                           label == "HAB" ~ "Biodiversity",
                           label == "ICO"~ "Sense of Place",
                           label == "SPP" ~"Biodiversity",
                           label == "LIV"~ "Coastal Livelihoods and Economies",
                           label == "LSP" ~ "Sense of Place",
                           label == "MAR" ~ "Food Provision",
                           label =="FIS" ~ "Food Provision",
                           !label %in% c("AO","CP", "CS", "ECO","HAB", "ICO","NP","SPP","LIV", "LSP","MAR", "TR", "FIS") ~ NA, #set as na if a lyer
                           TRUE ~ title)) |> #add goals column
  mutate(layer = ifelse(group == "layer",title, NA))

#create the plot

layers_plot <- visNetwork(nodes=nodes, edges=edges) %>% 
  visEdges(
    hoverWidth = 10,
    selectionWidth = 10,
    color = list(highlight = "#a7354e", hover = "#a7354e")
    ) |> #add highlight color 
   visInteraction(hover = TRUE) |> 
  visOptions(clickToUse = F, 
             highlightNearest = 
               list(enabled = TRUE, algorithm = "hierarchical",
                  degree = 1, hideColor = "rgba(0,0,0,0)", labelOnly = TRUE), #set color for lines when not selected 
                  nodesIdSelection = FALSE, #don't allow dropdown for all nodes
             selectedBy = list(variable = "goal", highlight = TRUE))|> #add dropdown list for selecting by goal
   visGroups(
    groupname = "goals",
    color = "#a7354e",
    size = 200,
    shape = "circularImage",
    font = list(size =0)
  ) %>% #set colors and font for all goals
  visGroups(
    groupname = "layers",
    color = "#377cbf",
    size = 55,
    font = list(size =0)) |> #set colors and fonts for all goals
 visPhysics(solver = "forceAtlas2Based", 
            forceAtlas2Based = list(gravitationalConstant = -350,
                                    damping = 1) 
    )

saveWidget(layers_plot, "/Users/adelheid/Documents/Work/NCEAS/repositories/OHI-website/content/images/layers_blog/layer_connections.html")
```

# 
